---
title: "p8105 - Homework 1"
author: "Jie Yu"
date: 2018-09-15
output: github_document
---

```{r setup, include = FALSE}
# Load the package tidyverse, without displaying the code and the result
library(tidyverse)
```

# Problem 1

The chunk below creates a data frame containing a random sample of size 10 from a uniform[0, 5] distribution, a logical vector indicating whether elements of the sample are greater than 2, a (length-10) character vector, and a (length-10) factor vector.

```{r p1_dataframe}
# Set seed to ensure reproducibility
set.seed(123)  

# Create a data frame of 4 variables
p1_df = tibble(
  unif_samp = runif(10, min = 0, max = 5),
  vec_logical = unif_samp >= 2, 
  vec_char = c("I", "am", "glad", "to", "be", "here", "Wishing", "you", "happiness", "always"),
  vec_factor = factor(c("male", "male", "female", "female", "male", "female", "female", "male", "female", "female"))
)

```

Below is the data frame:

```{r show_the_dataframe, echo = FALSE}
# Show the data frame, without displaying the code
p1_df
```



Take the mean of each variable in the above data frame.

* The mean of variable `unif_samp` is `r mean(p1_df$unif_samp)`
* The mean of variable `vec_logical` is `r mean(p1_df$vec_logical)`
* The mean of variable `vec_char` is `r mean(p1_df$vec_char)`
* The mean of variable `vec_factor` is `r mean(p1_df$vec_factor)`

We get the mean of variables `unif_samp` and `vec_logical`, but we does not get the mean of variables `vec_char` and `vec_factor`. That is because we cannot calculate the mean of character and factor variables.



The chunk below applies the `as.numeric` function to the logical, character, and factor variables.

```{r apply_function_as.numeric, eval = FALSE}
as.numeric(p1_df$vec_logical)
as.numeric(p1_df$vec_character)
as.numeric(p1_df$vec_factor)

```

The scripts above convert the logical, character, and factor variables from their original types to numeric. 



The chunk below converts the character variable from character to factor to numeric; similarly, converts the factor variable from factor to character to numeric. 

```{r convert_variables}
# Convert the character variable to facter, then to numeric
as.numeric(factor(p1_df$vec_character))   

# Convert the factor variable to character, then to numeric
as.numeric(as.character(p1_df$vec_factor))
                         

```

We find that the first conversion succeeds, and outputs a sequence of numbers corresponding to the alphabetical order of the original character variables.

However, the second conversion outputs a sequence of NA values. From this experiment, We learn that character variable cannot be converted from character to numeric directly. It should be first converted to factor, and then to numeric. 
